---
title: "Master File"
author: "Joanna Duran, Nikhil Gupta, Max Moro"
date: "`r Sys.time()`"
always_allow_html: yes
output:
 html_document:
   toc: true
   toc_float: true
   toc_depth: 3
 word_document:
   reference_docx: word_template.docx
params:
  output.var: 'y3'
  log.pred: TRUE
  algo.forward.caret: TRUE
  algo.backward.caret: TRUE
  algo.stepwise.caret: TRUE
  algo.LASSO.caret: TRUE
  algo.LARS.caret: TRUE
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman) #manages autoinstallation of packages
pacman::p_load(usdm,tidyverse,ggplot2,PerformanceAnalytics,MASS,glmnet,investr,ggiraph,ggiraphExtra
               ,DT,mosaic,usdm,caTools,onewaytests,scales
               ,bestNormalize,car,rlist,ggforce,leaps,"dplyr", "DT", "mosaic", "MASS", "usdm", "tidyverse", "ggplot2", "PerformanceAnalytics", "caTools", "glmnet", "caret","leaps","doParallel","lars","reshape2", "ggforce","rlist", "bestNormalize","car","htmltools","knitr")

options(width = 120)
```

# User Inputs

```{r User Inputs}
output.var = params$output.var 

transform.abs = FALSE
log.pred = params$log.pred
norm.pred = FALSE
algo.forward.caret = params$algo.forward.caret
algo.backward.caret = params$algo.backward.caret
algo.stepwise.caret = params$algo.stepwise.caret
algo.LASSO.caret = params$algo.LASSO.caret
algo.LARS.caret = params$algo.LARS.caret
message("Parameters used for training/prediction: ")
str(params)
```

```{r setup Labels}
# Setup Labels
output.var.tr = if (log.pred == TRUE)  paste0(output.var,'.log') else  output.var.tr = output.var
```

# Loading Data

```{r load}
feat  = read.csv('../../Data/features_highprec.csv')
labels = read.csv('../../Data/labels.csv')
predictors = names(dplyr::select(feat,-JobName))
data.ori = inner_join(feat,labels,by='JobName')
#data.ori = inner_join(feat,select_at(labels,c('JobName',output.var)),by='JobName')
```
 
# Data validation

```{r data}
cc  = complete.cases(data.ori)
data.notComplete = data.ori[! cc,]
data = data.ori[cc,] %>% select_at(c(predictors,output.var,'JobName'))
message('Original cases: ',nrow(data.ori))
message('Non-Complete cases: ',nrow(data.notComplete))
message('Complete cases: ',nrow(data))
summary(dplyr::select_at(data,c('JobName',output.var)))
```

# Output Variable

The Output Variable **`r output.var`** shows right skewness, so will proceed with a log transformation

## Histogram

```{r histogram_y, fig.height=3}
df=gather(select_at(data,output.var))
ggplot(df, aes(x=value)) + 
  geom_histogram(aes(y=..density..),bins = 50,fill='light blue') + 
  geom_density() 
  #stat_function(fun = dnorm, n = 100, args = list(mean = mean(df$value), sd = sd(df$value)))  

```

## QQPlot

```{r qqplot_y, fig.height=3}
ggplot(gather(select_at(data,output.var)), aes(sample=value)) + 
  stat_qq() + 
  facet_wrap(~key, scales = 'free',ncol=4)

```


## Trasformation of Output Variable from `r output.var` to `r output.var.tr`

```{r Trans Output, fig.height=4}
if(log.pred==TRUE) data[[output.var.tr]] = log(data[[output.var]],10) else
  data[[output.var.tr]] = data[[output.var]]
df=gather(select_at(data,c(output.var,output.var.tr)))
ggplot(df, aes(value)) + 
  geom_histogram(aes(y=..density..),bins = 50,fill='light blue') + 
  geom_density() + 
  # stat_function(fun = dnorm, n = 100, args = list(mean = mean(df$value), sd = sd(df$value)))  
  facet_wrap(~key, scales = 'free',ncol=2)

ggplot(gather(select_at(data,c(output.var,output.var.tr))), aes(sample=value)) + 
  stat_qq() + 
  facet_wrap(~key, scales = 'free',ncol=4)

```

## Best Normalizator `r output.var`

Normalization of **`r output.var`** using bestNormalize package.  (suggested orderNorm)
This is cool, but I think is too far for the objective of the project

```{r bestNormy, cache=T,fig.height=4}
t=bestNormalize::bestNormalize(data[[output.var]])
t
qqnorm(data[[output.var]])
qqnorm(predict(t))
```

**orderNorm()** is a rank-based procedure by which the values of a vector are mapped to their percentile, which is then mapped to the same percentile of the normal distribution. Without the presence of ties, this essentially guarantees that the transformation leads to a uniform distribution


# Predictors

All predictors show a **Fat-Tail** situation, where the two tails are very tall, and a low distribution around the mean.
The orderNorm transformation can help (see [Best Normalizator] section)

## Interesting Predictors

Histograms

```{r Histo Interest Pred, fig.height=6}
cols = c('x11','x18','stat98','x7','stat110')
df=gather(select_at(data,cols))
ggplot(df, aes(value)) + 
  geom_histogram(aes(y=..density..),bins = 50,fill='light blue') + 
  geom_density() + 
  # stat_function(fun = dnorm, n = 100, args = list(mean = mean(df$value), sd = sd(df$value)))  
  facet_wrap(~key, scales = 'free',ncol=3)

# ggplot(gather(select_at(data,cols)), aes(sample=value)) + 
#   stat_qq()+
#   facet_wrap(~key, scales = 'free',ncol=2)

lapply(select_at(data,cols),summary)
```

Scatter plot vs. output variable **`r output.var.tr`

```{r Scatters Interes Pred, fig.height=6}

d = gather(dplyr::select_at(data,c(cols,output.var.tr)),key=target,value=value,-!!output.var.tr)
ggplot(data=d, aes_string(x='value',y=output.var.tr)) + 
  geom_point(color='light green',alpha=0.5) + 
  geom_smooth() + 
  facet_wrap(~target, scales = 'free',ncol=3)
```

## All Predictors

### Histograms

All indicators have a strong indication of  **Fat-Tails**

```{r Histogram All, fig.height=70, fig.width=7, cache=TRUE}
df=gather(select_at(data,predictors))
ggplot(df, aes(value)) + 
  geom_histogram(aes(y=..density..),bins = 50,fill='light blue') + 
  geom_density() + 
  # stat_function(fun = dnorm, n = 100, args = list(mean = mean(df$value), sd = sd(df$value)))  
  facet_wrap(~key, scales = 'free',ncol=4)
```

# Correlations

## With Output Variable 

```{r CorrAll}
#chart.Correlation(select(data,-JobName),  pch=21)
t=as.data.frame(round(cor(dplyr::select(data,-one_of(output.var.tr,'JobName'))
                          ,select_at(data,output.var.tr)),4))  %>%
  rownames_to_column(var='variable') %>% filter(variable != !!output.var) %>% arrange(-y3.log)
#DT::datatable(t)
message("Top Positive")
kable(head(arrange(t,desc(y3.log)),20))
message("Top Negative")
kable(head(arrange(t,y3.log),20))
```

## Between All Variables

```{r correlation, fig.height=7}
#chart.Correlation(select(data,-JobName),  pch=21)
t=as.data.frame(round(cor(dplyr::select(data,-one_of('JobName'))),4))
#DT::datatable(t,options=list(scrollX=T))
message("Showing only 10 variables")
kable(t[1:10,1:10])
```
 
## Scatter Plots with Output Variable

Scatter plots with all predictors and the output variable (`r output.var.tr`)

```{r scatter_plots, fig.height=90, fig.width=7, cache=TRUE}
d = gather(dplyr::select_at(data,c(predictors,output.var.tr)),key=target,value=value,-!!output.var.tr)
ggplot(data=d, aes_string(x='value',y=output.var.tr)) + 
  geom_point(color='light blue',alpha=0.5) + 
  geom_smooth() + 
  facet_wrap(~target, scales = 'free',ncol=4)
```

## Multicollinearity - VIF

No Multicollinearity among predictors

Showing Top predictor by VIF Value

```{r vif, cache=TRUE}
vifDF = usdm::vif(select_at(data,predictors)) %>% arrange(desc(VIF))
head(vifDF,15)
```
 
# Feature Eng

- Square Root  transformation for **x18**

```{r FeatEng}
data.tr=data %>%
  mutate(x18.sqrt = sqrt(x18)) 
cols=c('x18','x18.sqrt')
```

## Comparing Pre and Post Transformation Density Plots


```{r Compare, fig.height=3}
# ggplot(gather(select_at(data.tr,cols)), aes(value)) + 
#   geom_histogram(aes(y=..density..),bins = 50,fill='light blue') + 
#   geom_density() + 
#   facet_wrap(~key, scales = 'free',ncol=4)

d = gather(dplyr::select_at(data.tr,c(cols,output.var.tr)),key=target,value=value,-!!output.var.tr)
ggplot(data=d, aes_string(x='value',y=output.var.tr)) + 
  geom_point(color='light blue',alpha=0.5) + 
  geom_smooth() + 
  facet_wrap(~target, scales = 'free',ncol=4)

```

```{r feat_eng_update}
#removing unwanted variables
data.tr=data.tr %>%
  dplyr::select_at(names(data.tr)[! names(data.tr) %in% c('x18','y3','JobName')])

data=data.tr
label.names=output.var.tr
```

# Modeling

##  Train Test Split

```{r Train Test Split}
data = data[sample(nrow(data)),] # randomly shuffle data
split = sample.split(data[,label.names], SplitRatio = 0.8)

data.train = subset(data, split == TRUE)
data.test = subset(data, split == FALSE)
```

## Common Functions
```{r CommonFunctions}
plot.diagnostics <-  function(model, train) {
  plot(model)
  
  residuals = resid(model) # Plotted above in plot(lm.out)
  r.standard = rstandard(model)
  r.student = rstudent(model)
  
  df = data.frame(x=predict(model,train),y=r.student)
  p=ggplot(data=df,aes(x=x,y=y)) +
    geom_point(color='blue',alpha=0.5,shape=20,size=2) +
    geom_hline(yintercept = 0,size=1)+
    ylab("Student Residuals") +
    xlab("Predicted Values")+
    ggtitle("Student Residual Plot")
  plot(p)
  
  df = data.frame(x=predict(model,train),y=r.standard)
  p=ggplot(data=df,aes(x=x,y=y)) +
    geom_point(color='blue',alpha=0.5,shape=20,size=2) +
    geom_hline(yintercept = c(-2,0,2),size=1)+
    ylab("Student Residuals") +
    xlab("Predicted Values")+
    ggtitle("Student Residual Plot")
  plot(p)
  # Histogram
  df=data.frame(r.student)
  p=ggplot(data=df,aes(r.student)) +
    geom_histogram(aes(y=..density..),bins = 50,fill='blue',alpha=0.6) + 
    stat_function(fun = dnorm, n = 100, args = list(mean = 0, sd = 1)) +
    ylab("Density")+
    xlab("Studentized Residuals")+
    ggtitle("Distribution of Studentized Residuals")
  plot(p)
  # http://www.stat.columbia.edu/~martin/W2024/R7.pdf
  # Influential plots
  inf.meas = influence.measures(model)
  # print (summary(inf.meas)) # too much data
  
  # Leverage plot
  lev = hat(model.matrix(model))
  df=tibble::rownames_to_column(as.data.frame(lev),'id')
  p=ggplot(data=df,aes(x=as.numeric(id),y=lev)) +
    geom_point(color='blue',alpha=0.5,shape=20,size=2) +
    ylab('Leverage - check') + 
    xlab('Index')
  plot(p)
  # Cook's Distance
  cd = cooks.distance(model)
  df=tibble::rownames_to_column(as.data.frame(cd),'id')
  p=ggplot(data=df,aes(x=as.numeric(id),y=cd)) +
    geom_point(color='blue',alpha=0.5,shape=20,size=2) +
    geom_text(data=filter(df,cd>15/nrow(train)),aes(label=id),check_overlap=T,size=3,vjust=-.5)+
    ylab('Cooks distances') + 
    geom_hline(yintercept = c(4/nrow(train),0),size=1)+
    xlab('Index')
  plot(p)
  print (paste("Number of data points that have Cook's D > 4/n: ", length(cd[cd > 4/nrow(train)]), sep = "")) 
  print (paste("Number of data points that have Cook's D > 1: ", length(cd[cd > 1]), sep = "")) 
  return(cd)
}

# function to set up random seeds
# Based on http://jaehyeon-kim.github.io/2015/05/Setup-Random-Seeds-on-Caret-Package.html 
setCaretSeeds <- function(method = "cv", numbers = 1, repeats = 1, tunes = NULL, seed = 1701) {
  #B is the number of resamples and integer vector of M (numbers + tune length if any)
  B <- if (method == "cv") numbers
  else if(method == "repeatedcv") numbers * repeats
  else NULL
  if(is.null(length)) {
    seeds <- NULL
  } else {
    set.seed(seed = seed)
    seeds <- vector(mode = "list", length = B)
    seeds <- lapply(seeds, function(x) sample.int(n = 1000000
                                                  , size = numbers + ifelse(is.null(tunes), 0, tunes)))
    seeds[[length(seeds) + 1]] <- sample.int(n = 1000000, size = 1)
  }
  # return seeds
  seeds
}

train.caret.glmselect = function(formula, data, method
                                 ,subopt = NULL, feature.names
                                 , train.control = NULL, tune.grid = NULL, pre.proc = NULL){
  
  if(is.null(train.control)){
    train.control <- trainControl(method = "cv"
                              ,number = 10
                              ,seeds = setCaretSeeds(method = "cv"
                                                     , numbers = 10
                                                     , seed = 1701)
                              ,search = "grid"
                              ,verboseIter = TRUE
                              ,allowParallel = TRUE
                              )
  }
  
  if(is.null(tune.grid)){
    if (method == 'leapForward' | method == 'leapBackward' | method == 'leapSeq'){
      tune.grid = data.frame(nvmax = 1:length(feature.names))
    }
    if (method == 'glmnet' && subopt == 'LASSO'){
      # Will only show 1 Lambda value during training, but that is OK
      # https://stackoverflow.com/questions/47526544/why-need-to-tune-lambda-with-carettrain-method-glmnet-and-cv-glmnet
      # Another option for LASSO is this: https://github.com/topepo/caret/blob/master/RegressionTests/Code/lasso.R
      lambda = 10^seq(-2,0, length =100)
      alpha = c(1)
      tune.grid = expand.grid(alpha = alpha,lambda = lambda)
    }
    if (method == 'lars'){
      # https://github.com/topepo/caret/blob/master/RegressionTests/Code/lars.R
      fraction = seq(0, 1, length = 100)
      tune.grid = expand.grid(fraction = fraction)
      pre.proc = c("center", "scale") 
    }
  }
  
  # http://sshaikh.org/2015/05/06/parallelize-machine-learning-in-r-with-multi-core-cpus/
  cl <- makeCluster(ceiling(detectCores()*0.85)) # use 75% of cores only, leave rest for other tasks
  registerDoParallel(cl)

  set.seed(1) 
  # note that the seed has to actually be set just before this function is called
  # settign is above just not ensure reproducibility for some reason
  model.caret <- caret::train(formula
                              , data = data
                              , method = method
                              , tuneGrid = tune.grid
                              , trControl = train.control
                              , preProc = pre.proc
                              )
  
  stopCluster(cl)
  registerDoSEQ() # register sequential engine in case you are not using this function anymore
  
  if (method == 'leapForward' | method == 'leapBackward' | method == 'leapSeq'){
    print("All models results")
    print(model.caret$results) # all model results
    print("Best Model")
    print(model.caret$bestTune) # best model
    model = model.caret$finalModel

    # Metrics Plot 
    dataPlot = model.caret$results %>%
      gather(key='metric',value='value',-nvmax) %>%
      dplyr::filter(metric %in% c('MAE','RMSE','Rsquared'))
    metricsPlot = ggplot(data=dataPlot,aes(x=nvmax,y=value) ) +
      geom_line(color='lightblue4') +
      geom_point(color='blue',alpha=0.7,size=.9) +
      facet_wrap(~metric,ncol=2,scales='free_y')+
      theme_light()
    plot(metricsPlot)
    
    # Residuals Plot
    # leap function does not support studentized residuals
    dataPlot=data.frame(pred=predict(model.caret,data),res=resid(model.caret))
    residPlot = ggplot(dataPlot,aes(x=pred,y=res)) +
      geom_point(color='light blue',alpha=0.7) +
      geom_smooth(method="lm")+
      theme_light()
    plot(residPlot)
   
    residHistogram = ggplot(dataPlot,aes(x=res)) +
      geom_histogram(aes(y=..density..),fill='light blue',alpha=1) +
      #geom_density(color='lightblue4') + 
      stat_function(fun = dnorm, n = 100, args = list(mean = mean(dataPlot$res)
                                                       , sd = sd(dataPlot$res)),color='lightblue4')  
      theme_light()
    plot(residHistogram)
    id = rownames(model.caret$bestTune)    
    # Provides the coefficients of the best model
    # regsubsets doens return a full model (see documentation of regsubset), so we need to recalcualte themodel
    # https://stackoverflow.com/questions/13063762/how-to-obtain-a-lm-object-from-regsubsets
    print("Coefficients of final model:")
    coefs <- coef(model, id=id)
    #calculate the model to the the coef intervals
    nams <- names(coefs)
    nams <- nams[!nams %in% "(Intercept)"]
    response <-  as.character(formula[[2]])
    form <- as.formula(paste(response, paste(nams, collapse = " + "), sep = " ~ "))
    mod <- lm(form, data = data)
    #coefs
    #coef(mod)
    print(car::Confint(mod))
    return(list(model = model,id = id, residPlot = residPlot, residHistogram=residHistogram
                ,modelLM=mod))
  }
  if (method == 'glmnet' && subopt == 'LASSO'){
    print(model.caret)
    print(plot(model.caret))
    print(model.caret$bestTune)
    
    print(model.caret$results)
    model=model.caret$finalModel
    # Metrics Plot 
    dataPlot = model.caret$results %>%
      gather(key='metric',value='value',-lambda) %>%
      dplyr::filter(metric %in% c('MAE','RMSE','Rsquared'))
    metricsPlot = ggplot(data=dataPlot,aes(x=lambda,y=value) ) +
      geom_line(color='lightblue4') +
      geom_point(color='blue',alpha=0.7,size=.9) +
      facet_wrap(~metric,ncol=2,scales='free_y')+
      theme_light()
    plot(metricsPlot)
    
    # Residuals Plot 
    dataPlot=data.frame(pred=predict(model.caret,data),res=resid(model.caret))
    residPlot = ggplot(dataPlot,aes(x=pred,y=res)) +
      geom_point(color='light blue',alpha=0.7) +
      geom_smooth(method="lm")+
      theme_light()
    plot(residPlot)

    residHistogram = ggplot(dataPlot,aes(x=res)) +
      geom_histogram(aes(y=..density..),fill='light blue',alpha=1) +
      #geom_density(color='lightblue4') +
      stat_function(fun = dnorm, n = 100, args = list(mean = mean(dataPlot$res)
                                                       , sd = sd(dataPlot$res)),color='lightblue4')  
      theme_light()
    plot(residHistogram)
    
    print("Coefficients") 
    #no interval for glmnet: https://stackoverflow.com/questions/39750965/confidence-intervals-for-ridge-regression
    t=coef(model,s=model.caret$bestTune$lambda)
    model.coef = t[which(t[,1]!=0),]
    print(as.data.frame(model.coef))
    id = NULL # not really needed but added for consistency
    return(list(model = model.caret,id = id, residPlot = residPlot, metricsPlot=metricsPlot ))
  }
  if (method == 'lars'){
    print(model.caret)
    print(plot(model.caret))
    print(model.caret$bestTune)
    
    # Metrics Plot
    dataPlot = model.caret$results %>%
        gather(key='metric',value='value',-fraction) %>%
      dplyr::filter(metric %in% c('MAE','RMSE','Rsquared'))
    metricsPlot = ggplot(data=dataPlot,aes(x=fraction,y=value) ) +
      geom_line(color='lightblue4') +
      geom_point(color='blue',alpha=0.7,size=.9) +
      facet_wrap(~metric,ncol=2,scales='free_y')+
      theme_light()
    plot(metricsPlot)
    
    # Residuals Plot
    dataPlot=data.frame(pred=predict(model.caret,data),res=resid(model.caret))
    residPlot = ggplot(dataPlot,aes(x=pred,y=res)) +
      geom_point(color='light blue',alpha=0.7) +
      geom_smooth(method="lm")+
      theme_light()
    plot(residPlot)

    residHistogram = ggplot(dataPlot,aes(x=res)) +
      geom_histogram(aes(y=..density..),fill='light blue',alpha=1) +
      #geom_density(color='lightblue4') + 
      stat_function(fun = dnorm, n = 100, args = list(mean = mean(dataPlot$res)
                                                       , sd = sd(dataPlot$res)),color='lightblue4')  
      theme_light()
    plot(residHistogram)
    
    print("Coefficients") 
    t=coef(model.caret$finalModel,s=model.caret$bestTune$fraction,mode='fraction')
    model.coef = t[which(t!=0)]
    print(model.coef)
    id = NULL # not really needed but added for consistency
    return(list(model = model.caret,id = id, residPlot = residPlot, residHistogram=residHistogram))
  }
}

# https://stackoverflow.com/questions/48265743/linear-model-subset-selection-goodness-of-fit-with-k-fold-cross-validation
# changed slightly since call[[2]] was just returning "formula" without actually returnign the value in formula
predict.regsubsets <- function(object, newdata, id, formula, ...) {
    #form <- as.formula(object$call[[2]])
    mat <- model.matrix(formula, newdata) # adds intercept and expands any interaction terms
    coefi <- coef(object, id = id)
    xvars <- names(coefi)
    return(mat[,xvars]%*%coefi)
}
  
test.model = function(model, test, level=0.95
                      ,draw.limits = FALSE, good = 0.1, ok = 0.15
                      ,method = NULL, subopt = NULL
                      ,id = NULL, formula, feature.names, label.names
                      ,transformation = NULL){
  ## if using caret for glm select equivalent functionality, 
  ## need to pass formula (full is ok as it will select subset of variables from there)
  if (is.null(method)){
    pred = predict(model, newdata=test, interval="confidence", level = level) 
  }
  
  if (method == 'leapForward' | method == 'leapBackward' | method == 'leapSeq'){
    pred = predict.regsubsets(model, newdata = test, id = id, formula = formula)
  }
  
  if (method == 'glmnet' && subopt == 'LASSO'){
    xtest = as.matrix(test[,feature.names]) 
    pred=as.data.frame(predict(model, xtest))
  }
  
  if (method == 'lars'){
    pred=as.data.frame(predict(model, newdata = test))
  }
    
  # Summary of predicted values
  print ("Summary of predicted values: ")
  print(summary(pred[,1]))

  test.mse = mean((test[,label.names]-pred[,1])^2)
  print (paste(method, subopt, "Test MSE:", test.mse, sep=" "))
  
  if(log.pred == TRUE || norm.pred == TRUE){
    # plot transformewd comparison first
    df=data.frame(x=test[,label.names],y=pred[,1])
    ggplot(df,aes(x=x,y=y)) +
      geom_point(color='blue',alpha=0.5,shape=20,size=2) +
      geom_abline(slope=1,intercept=0,color='black',size=1) +
      #scale_y_continuous(limits=c(min(df),max(df)))+
      xlab("Actual (Transformed)")+
      ylab("Predicted (Transformed)")
  }
    
  if (log.pred == FALSE && norm.pred == FALSE){
    x = test[,label.names]
    y = pred[,1]
  }
  if (log.pred == TRUE){
    x = 10^test[,label.names]
    y = 10^pred[,1]  
  }
  if (norm.pred == TRUE){
    x = predict(transformation, test[,label.names], inverse = TRUE)
    y = predict(transformation, pred[,1], inverse = TRUE)
  }

  df=data.frame(x,y)
  ggplot(df,aes(x,y)) +
    geom_point(color='blue',alpha=0.5,shape=20,size=2) +
    geom_abline(slope=c(1+good,1-good,1+ok,1-ok)
                ,intercept=rep(0,4),color=c('dark green','dark green','dark red','dark red'),size=1,alpha=0.8) +
    #scale_y_continuous(limits=c(min(df),max(df)))+
    xlab("Actual")+
    ylab("Predicted") 
    
 
}
```

## Setup Formulae
```{r Formulae}
n <- names(data.train)
 formula <- as.formula(paste(paste(n[n %in% label.names], collapse = " + ")
                             ," ~", paste(n[!n %in% label.names], collapse = " + "))) 

grand.mean.formula = as.formula(paste(paste(n[n %in% label.names], collapse = " + ")," ~ 1"))

print(formula)
print(grand.mean.formula)

# Update feature.names because we may have transformed some features
feature.names = n[!n %in% label.names]
```

## Full Model

```{r Full Model}
model.full = lm(formula , data.train)
summary(model.full)
cd.full = plot.diagnostics(model=model.full, train=data.train)
```

## Checking with removal of high influence points
```{r HighInfluenceCheck, cache=T}
high.cd = names(cd.full[cd.full > 4/nrow(data.train)])

#save dataset with high.cd flagged
t = data.train %>% 
  rownames_to_column() %>%
  mutate(high.cd = ifelse(rowname %in% high.cd,1,0))
write.csv(t,file='data_high_cd_flag.csv',row.names = F)
###
data.train2 = data.train[!(rownames(data.train)) %in% high.cd,]
model.full2 = lm(formula , data.train2)
summary(model.full2)
cd.full2 = plot.diagnostics(model.full2, data.train2)

# much more normal residuals than before. 
# Checking to see if distributions are different and if so whcih variables
# High Leverage Plot 
plotData = data.train %>% 
  rownames_to_column() %>%
  mutate(type=ifelse(rowname %in% high.cd,'High','Normal')) %>%
  dplyr::select(type,target=one_of(label.names))

ggplot(data=plotData, aes(x=type,y=target)) +
  geom_boxplot(fill='light blue',outlier.shape=NA) +
  scale_y_continuous(name="Target Variable Values",label=scales::comma_format(accuracy=.1)) +
  theme_light() +
  ggtitle('Distribution of High Leverage Points and Normal  Points')


```

```{r 2SampleTest}
# 2 sample t-tests

plotData = data.train %>% 
  rownames_to_column() %>%
  mutate(type=ifelse(rowname %in% high.cd,'High','Normal')) %>%
  dplyr::select(type,one_of(feature.names))

comp.test = lapply(dplyr::select(plotData, one_of(feature.names))
                   , function(x) t.test(x ~ plotData$type, var.equal = TRUE)) 

sig.comp = list.filter(comp.test, p.value < 0.05)
sapply(sig.comp, function(x) x[['p.value']])

mm = melt(plotData, id=c('type')) %>% filter(variable %in% names(sig.comp))

ggplot(mm,aes(x=type, y=value)) +
  geom_boxplot()+
  facet_wrap(~variable, ncol=5, scales = 'free_y') +
  scale_y_continuous(name="values",label=scales::comma_format(accuracy=.1)) +
  ggtitle('Distribution of High Leverage Points and Normal Points')

```


```{r High Leverage plot, fig.height=40, cache=T}
# Distribution (box) Plots
mm = melt(plotData, id=c('type'))

ggplot(mm,aes(x=type, y=value)) +
  geom_boxplot()+
  facet_wrap(~variable, ncol=8, scales = 'free_y') +
  scale_y_continuous(name="values",label=scales::comma_format(accuracy=.1)) +
  ggtitle('Distribution of High Leverage Points and Normal Points')

```

## Grand Means Model
```{r Grand Mean Model}
model.null = lm(grand.mean.formula, data.train)
summary(model.null)
```

## Variable Selection 
Basic: http://www.stat.columbia.edu/~martin/W2024/R10.pdf 
Cross Validation + Other Metrics: http://www.sthda.com/english/articles/37-model-selection-essentials-in-r/154-stepwise-regression-essentials-in-r/

### Forward Selection with CV  

#### Train
```{r Forward,fig.height=5}
if (algo.forward.caret == TRUE){
  set.seed(1)
  returned = train.caret.glmselect(formula = formula
                                   , data = data.train
                                   , method = "leapForward"
                                   , feature.names = feature.names)
  model.forward = returned$model
  id = returned$id
}
```

#### Test
```{r Forward Test, fig.height=5}
if (algo.forward.caret == TRUE){
    test.model(model=model.forward, test=data.test
             ,method = 'leapForward',subopt = NULL
             ,formula = formula, feature.names = feature.names, label.names = label.names
             ,id = id
             ,draw.limits = TRUE, transformation = t)
}

```

### Backward Elimination with CV 
#### Train
```{r Backward, fig.height=5}
if (algo.backward.caret == TRUE){
  set.seed(1)
  returned = train.caret.glmselect(formula = formula
                                   ,data =  data.train
                                   ,method = "leapBackward"
                                   ,feature.names =  feature.names)
  model.backward = returned$model
  id = returned$id
}
```

#### Test
```{r Backward Test,fig.height=5}
if (algo.backward.caret == TRUE){
  test.model(model.backward, data.test
             ,method = 'leapBackward',subopt = NULL
             ,formula = formula, feature.names = feature.names, label.names = label.names
             ,id = id
             ,draw.limits = TRUE, transformation = t)
}
```

### Stepwise Selection with CV 
#### Train
```{r Stepwise,fig.height=5,cache=T}
if (algo.stepwise.caret == TRUE){
  set.seed(1)
  returned = train.caret.glmselect(formula = formula
                                   ,data =  data.train
                                   ,method = "leapSeq"
                                   ,feature.names = feature.names)
  model.stepwise = returned$model
  id = returned$id
}
```

#### Test
```{r Stepwise Test,fig.height=5}
if (algo.stepwise.caret == TRUE){
  test.model(model.stepwise, data.test
             ,method = 'leapSeq',subopt = NULL
             ,formula = formula, feature.names = feature.names, label.names = label.names
             ,id = id
             ,draw.limits = TRUE, transformation = t)
  
}
```

### LASSO with CV 

#### Train
```{r Lasso, fig.height=5}
if (algo.LASSO.caret == TRUE){
  set.seed(1)
  tune.grid= expand.grid(alpha = 1,lambda = 10^seq(from=-4,to=-2,length=100))
  returned = train.caret.glmselect(formula = formula
                                   ,data =  data.train
                                   ,method = "glmnet"
                                   ,subopt = 'LASSO'
                                   ,tune.grid = tune.grid
                                   ,feature.names = feature.names)
  model.LASSO.caret = returned$model
}
```


#### Test
```{r Lasso Test}
if (algo.LASSO.caret == TRUE){
  test.model(model.LASSO.caret, data.test
             ,method = 'glmnet',subopt = "LASSO"
             ,formula = formula, feature.names = feature.names, label.names = label.names
             ,draw.limits = TRUE, transformation = t)
}
```

### LARS with CV  
#### Train
```{r LARS, fig.height=5}
if (algo.LARS.caret == TRUE){
  set.seed(1)
  returned = train.caret.glmselect(formula = formula
                                   ,data =  data.train
                                   ,method = "lars"
                                   ,subopt = 'NULL'
                                   ,feature.names = feature.names)
  model.LARS.caret = returned$model
}
```


#### Test
```{r LARS Test, fig.height=5}
if (algo.LARS.caret == TRUE){
  test.model(model.LARS.caret, data.test
             ,method = 'lars',subopt = NULL
             ,formula = formula, feature.names = feature.names, label.names = label.names
             ,draw.limits = TRUE, transformation = t)
}
```



# Session Info

```{r sessionInfo}
sessionInfo()
```