---
title: "Master File"
author: "Max Moro"
date: "`r Sys.time()`"
output:
 html_document:
   toc: true
   toc_float: true
   toc_depth: 3
params:
  output.var: 'y3'
  log.pred: TRUE
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(pacman) #manages autoinstallation of packages
pacman::p_load(usdm,tidyverse,ggplot2,PerformanceAnalytics,MASS,glmnet,investr,ggiraph,ggiraphExtra
               ,DT,mosaic,usdm,caTools,onewaytests
               ,bestNormalize)
sessionInfo()
```

# User Inputs

```{r User Inputs}
output.var = params$output.var 
transform.abs = params$transform.abs
log.pred = params$log.pred
message("Parameters used for training/prediction: ")
str(params)
```


```{r}
# Setup Labels
# alt.scale.label.name = Alternate Scale variable name
#   - if predicting on log, then alt.scale is normal scale
#   - if predicting on normal scale, then alt.scale is log scale
if (log.pred == TRUE){
  output.var.tr = paste0(output.var,'.log')
}
if (log.pred == FALSE){
  output.var.tr = output.var
}
```

# Loading Data

```{r load}
feat  = read.csv('../../Data/features_highprec.csv')
labels = read.csv('../../Data/labels.csv')
predictors = names(dplyr::select(feat,-JobName))
data.ori = inner_join(feat,labels,by='JobName')
#data.ori = inner_join(feat,select_at(labels,c('JobName',output.var)),by='JobName')
```
 
# Data validation

```{r data}
cc  = complete.cases(data.ori)
data.notComplete = data.ori[! cc,]
data = data.ori[cc,] %>% select_at(c(predictors,output.var,'JobName'))
message('Original cases: ',nrow(data.ori))
message('Non-Complete cases: ',nrow(data.notComplete))
message('Complete cases: ',nrow(data))
```

# Output Variable

The Output Variable **`r output.var`** shows  right skewness, so will proceed with a log transofrmation

## Histogram

```{r fig.height=3}
ggplot(gather(select_at(data,output.var)), aes(value)) + 
  geom_histogram(aes(y=..density..),bins = 50,fill='light blue') + 
  geom_density() + 
  facet_wrap(~key, scales = 'free',ncol=4)

```

## QQPlot

```{r fig.height=3}
ggplot(gather(select_at(data,output.var)), aes(sample=value)) + 
  stat_qq() + 
  facet_wrap(~key, scales = 'free',ncol=4)

```

## Best Normalizator `r output.var`

Normalization of **`r output.var`** using bestNormalize package.  (suggested orderNorm)
This is cool, but I think is too far for the objective of the project

```{r bestNormy, cache=T,fig.height=4}
t=bestNormalize::bestNormalize(data[[output.var]])
t
qqnorm(data[[output.var]])
qqnorm(predict(t))
```

**orderNorm()** is a rank-based procedure by which the values of a vector are mapped to their percentile, which is then mapped to the same percentile of the normal distribution. Without the presence of ties, this essentially guarantees that the transformation leads to a uniform distribution


## Trasformation of Output Variable from `r output.var` to `r output.var.tr`

```{r}
if(log.pred==TRUE) data[[output.var.tr]] = log(data[[output.var]],10) else
  data[[output.var.tr]] = data[[output.var]]

ggplot(gather(select_at(data,c(output.var,output.var.tr))), aes(value)) + 
  geom_histogram(aes(y=..density..),bins = 50,fill='light blue') + 
  geom_density() + 
  facet_wrap(~key, scales = 'free',ncol=4)

```

  
# Predictors

All predictors show a **Fat-Tail** situation, where the two tails are very tall, and a low distribution around the mean.
The orderNorm transromation can help (see [Best Normalizator] section)

## Interesting Predictors

Histogram and QQ plot

```{r fig.height=9}
cols = c('x11','x18','stat98','x7','stat110')
ggplot(gather(select_at(data,cols)), aes(value)) + 
  geom_histogram(aes(y=..density..),bins = 50,fill='light blue') + 
  geom_density() + 
  facet_wrap(~key, scales = 'free',ncol=2)

ggplot(gather(select_at(data,cols)), aes(sample=value)) + 
  stat_qq()+
  facet_wrap(~key, scales = 'free',ncol=2)

lapply(select_at(data,cols),summary)
```

Scatter plot vs. output variable **`r output.var.tr`

```{r fig.height=6}

d = gather(dplyr::select_at(data,c(cols,output.var.tr)),key=target,value=value,-!!output.var.tr)
ggplot(data=d, aes_string(x='value',y=output.var.tr)) + 
  geom_point(color='light green',alpha=0.5) + 
  geom_smooth() + 
  facet_wrap(~target, scales = 'free',ncol=2)
```


## Histograms

All indicators have a strong indication of  **Fat-Tails**

```{r density, fig.height=70, fig.width=7, cache=TRUE}
ggplot(gather(select_at(data,predictors)), aes(value)) + 
  geom_histogram(aes(y=..density..),bins = 50,fill='light blue') + 
  geom_density() + 
  facet_wrap(~key, scales = 'free',ncol=4)
```

## QQPlots

```{r qqplot, fig.height=70, fig.width=7, cache=TRUE}
ggplot(gather(select_at(data,predictors)), aes(sample=value)) + 
  stat_qq() + 
  facet_wrap(~key, scales = 'free',ncol=4)
```


# Correlations

## With Output Variable 

```{r}
#chart.Correlation(select(data,-JobName),  pch=21)
t=round(cor(dplyr::select(data,-one_of(output.var.tr,'JobName')),select_at(data,output.var.tr)),4)
DT::datatable(t)
```

## All  Variables

```{r correlation, fig.height=7}
#chart.Correlation(select(data,-JobName),  pch=21)
t=round(cor(dplyr::select(data,-one_of('JobName'))),4)
DT::datatable(t,options=list(scrollX=T))
```
 
## Scatter Plots with Output Variable

Scatter plots with all predictors and the output variable (`r output.var.tr`)

```{r scatter_plots, fig.height=90, fig.width=7, cache=TRUE}
d = gather(dplyr::select_at(data,c(predictors,output.var.tr)),key=target,value=value,-!!output.var.tr)
ggplot(data=d, aes_string(x='value',y=output.var.tr)) + 
  geom_point(color='light blue',alpha=0.5) + 
  geom_smooth() + 
  facet_wrap(~target, scales = 'free',ncol=4)
```

## Multicollinearity - VIF

No Multicollinearity among predictors

Showing Top predictor by VIF Value

```{r vif, cache=TRUE}
vifDF = usdm::vif(select_at(data,predictors)) %>% arrange(desc(VIF))
head(vifDF,10)
```
 
# Feature Eng

- No trasnformation for x18

- log transformatio for `r output.var`

```{r}
data.tr=data %>%
  mutate(x18.sqrt = sqrt(x18)) 
cols=c('x18','x18.sqrt')
```

## Comparing Pre and Post Transformation Density Plots


```{r fig.height=3}
ggplot(gather(select_at(data.tr,cols)), aes(value)) + 
  geom_histogram(aes(y=..density..),bins = 50,fill='light blue') + 
  geom_density() + 
  facet_wrap(~key, scales = 'free',ncol=4)
```


```{r fig.height=3}
d = gather(dplyr::select_at(data.tr,c(cols,output.var.tr)),key=target,value=value,-!!output.var.tr)
ggplot(data=d, aes_string(x='value',y=output.var.tr)) + 
  geom_point(color='light blue',alpha=0.5) + 
  geom_smooth() + 
  facet_wrap(~target, scales = 'free',ncol=4)
```

```{r}
#removing unwanted variables
data.tr=data.tr %>%
  dplyr::select_at(names(data.tr)[! names(data.tr) %in% c('x18sqrt','y3')])
```

# Conclusion

- the target ariable **`r output.var`** can be LOG transformed

- the predictor x18 is not improving with SQR trasformatioatn

- all predictors could benefit with a *orderNorm* transformation